---
title: "Data Import Lecture"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

## Import some data

I want to import `FAS_litters.csv`.

```{r}
litters_df = read_csv("data/FAS_litters.csv") #relative path
```

I imported the dataset. Now I want better variable names.
```{r}
names(litters_df)

litters_df = janitor::clean_names(litters_df)
```

Now I have better names. Let's look at the dataset.

```{r} 
#printing the tibble: shows first 10 rows, how many more rows, variables, variable types, highlights NAs in red
litters_df 

head(litters_df, 20) # print the first 20 rows
tail(litters_df) # should check there are no empty rows after importing
```

A useful function to pull up the entire dataset in a new window, not output:

```{r, eval = FALSE}
view(litters_df) # make sure you do eval = FALSE or do this in the command line
```

'skimr' tells you how many rows, columns, how many character vs. numeric vs. factors, how many variables are complete or missing rate, min and max group size, how many unique values, prints histogram of numeric variables and basic descriptive stats

```{r}
skimr::skim(litters_df)
```






