---
title: "Data Import Lecture"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Import some data

I want to import `FAS_litters.csv`.

```{r}
litters_df = read_csv("data/FAS_litters.csv") #relative path
```

I imported the dataset. Now I want better variable names.
```{r}
names(litters_df)

litters_df = janitor::clean_names(litters_df)
```

Now I have better names. Let's look at the dataset.

```{r} 
#printing the tibble: shows first 10 rows, how many more rows, variables, variable types, highlights NAs in red
litters_df 

head(litters_df, 20) # print the first 20 rows
tail(litters_df) # should check there are no empty rows after importing
```

A useful function to pull up the entire dataset in a new window, not output:

```{r, eval = FALSE}
view(litters_df) # make sure you do eval = FALSE or do this in the command line
```

'skimr' tells you how many rows, columns, how many character vs. numeric vs. factors, how many variables are complete or missing rate, min and max group size, how many unique values, prints histogram of numeric variables and basic descriptive stats

```{r}
skimr::skim(litters_df)
```

## Arguments in `read_csv`

```{r}
litters_df = read_csv(
  "data/FAS_litters.csv", 
  skip = 5,   # useful if there are extra notes in some top rows
  col_names = FALSE,
  na = "Low8") # tell it that this should be missing and convert it to NA (making this up)
```

## Parsing Columns

If you want to declare column types:

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)
```

Or to change just one column type, specify one at a time:

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(`GD of Birth` = col_integer())
)
```




## Reading from Excel

Use `?read_excel` to see the help file for a function

Reading MLB Data:

```{r}
mlb11_df = read_excel("data/mlb11.xlsx")
```

can specify sheet or range

LotR Words data is next - but it's only in certain cells

```{r}
fellow_df = read_excel("data/LotR_Words.xlsx", range = "B3:D6")
```



## Read a SAS file

Comes from Haven package.
There are also ways to read Stata, SPSS, MatLab, etc.

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```


## Why to never use `read.csv`:

* It imports as a dataframe instead of a tibble. Tibbles print nicer.


```{r}
litters_df_base = read.csv("data/FAS_litters.csv")
```


## Exporting Data

use `write_csv`

```{r}
write_csv(fellow_df, "data/fellowship_words.csv")
```

