---
title: "Data Manipulation Lecture"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

## Import some data

Importing `FAS_litters.csv` and  clean up the variable names.

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_df = read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## `select` function

Select only the columns you're interested in.

```{r}
select(litters_df, group, litter_number) # selects these variables

select(litters_df, group, gd0_weight, gd18_weight) 

select(litters_df, group, gd0_weight:gd_of_birth) # selects everything in range

select(litters_df, -litter_number) # selects everything but litter_num

select(litters_df, group) # selects the one variable, output = data frame

#pull(litters_df, group) # selects the data from this one variable
```

Renaming Variables

```{r}
select(litters_df, group_new = group, litter_number) # renames a variable

rename(litters_df, group_new2 = group) # another way to change column name
```

Relocating variables 

```{r}
select(litters_df, litter_number, everything()) # reorders just one, keeping everything else

relocate(litters_df, litter_number) # alternate way
```

Learning assessment:
In the pups data, select the columns containing litter number, sex, and PD ears.

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter` rows

Getting rid of rows - filtering by characteristic

```{r}
filter(litters_df, gd_of_birth == 20) # selecting if equal to a value
filter(litters_df, group == "Con7")

filter(litters_df, gd0_weight < 23) # selecting using comparison

filter(litters_df, pups_survive != 4) # all litters with not 4 surviving

filter(litters_df, !(group == "Con7"))

filter(litters_df, group %in% c("Con7", "Con8")) # filtering using list

filter(litters_df, group == "Con7", gd_of_birth == 20) # multiple conditions: AND
filter(litters_df, group == "Con7" & gd_of_birth == 20) # multiple conditions: AND

filter(litters_df, group == "Con7" | gd_of_birth == 20) # multiple conditions: OR
```

If you want to filter based on missingness, it's easier to the use drop_na

```{r}
drop_na(litters_df) # gets rid of rows with ANY missing data

drop_na(litters_df, gd0_weight) # gets rid of rows only missing this variable
```

Learning Assessment in the pups data:
1. filter to include only pups with sex 1
2. filter to include only pups with PD walk less than 11 and sex 2

```{r}
filter(pups_df, sex == 1)

filter(pups_df, pd_walk < 11, sex == 2)
```

## `mutate`

Adding or changing columns

```{r}
mutate(litters_df, 
  weight_change = gd18_weight - gd0_weight,
  group = str_to_lower(group)) # makes everything lowercase
```

## `arrange`

Rearranging the data (reordering observations).

```{r}
arrange(litters_df, gd0_weight) # order observations by ascending gd0_weight

arrange(litters_df, desc(gd0_weight)) # in descending order

arrange(litters_df, gd_of_birth, gd0_weight) # arranging by multiple things
```












