---
title: "Tidy Data Lecture"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## pivot longer

Always use `pivot_longer()` and not `gather()`.

Load the PULSE data.

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Data is currently in wide format - columns for baseline, 1 mo, 6 mo, 12 mo, etc.

Using pivot function

```{r}
pulse_tidy = 
  pulse_df %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_", # removes prefix from beginning of the var name
    values_to = "bdi"
  ) %>% 
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit) # converting visit variable to a factor
  )
```

## pivot_wider

going from long to wide data format.

always use `pivot_wider()`, not `spread()`.

making up a results data table

```{r}
analysis_df = 
  tibble(
    group = c("treatment", "treatment", "control", "control"),
    time = c("a", "b", "a", "b"),
    group_mean = c(4, 8, 3, 6) 
  )

analysis_df %>% 
  pivot_wider(
    names_from = "time",
    values_from = "group_mean"
  ) %>% 
  knitr::kable() # function to make printing tables prettier
```



## bind_rows

Like appending.

Always use `bind_rows` and not `rbind`.

Importing the LotR movie words data file

```{r}
fellowship_df = 
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_rings")

two_towers_df = 
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_df = 
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")

lotr_df = 
  bind_rows(fellowship_df, two_towers_df, return_df) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words"
  ) %>% 
  mutate(race = str_to_lower(race)) %>% 
  relocate(movie)
```



## joins

Imports and cleans litters and pups data.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group, into = c("dose", "day_of_tx"), 3) %>% # splitting into 2 variables
  relocate(litter_number) %>% 
  mutate(dose = str_to_lower(dose))
  
pups_df = 
  read_csv("data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))
```

Joining pups and litters data.

```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") %>% 
  relocate(litter_number, dose, day_of_tx)
```





